---
title: "Week 9 Case Study"
author: "Tyson Brost"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(readxl)
library(mosaic)
library(pander)
library(lubridate)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
sales_dat <- read.csv("https://byuistats.github.io/M335/data/sales.csv")
```

## Background

Read in the data from https://byuistats.github.io/M335/data/sales.csv and format it for visualization and analysis.

The data are for businesses in the mountain time zone make sure you read in times correctly.

This is point of sale (pos) data, so you will need to use library(lubridate) to create the correct time aggregations.

Check the data for any inaccuracies.

Help your boss understand which business is the best investment through visualizations.
Provide an understanding and recommendation for hours of operation.
We don’t have employee numbers, but sales traffic can help. Provide some visualizations on customer traffic.
Provide a final comparison of the six companies and a final recommendation.

#### Pander doesn't display properly in the .md. use the html for best viewing

## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
sales_dat$Time <- ymd_hms(sales_dat$Time, tz = "MST")
sales_dat$Time <- ceiling_date(sales_dat$Time, "hour")
sales_dat$hour <- format(as.POSIXct(sales_dat$Time), format = "%H")
sales_dat$day <- weekdays(as.POSIXct(sales_dat$Time), abbreviate = F)
sales_dat$week <- week(as.POSIXct(sales_dat$Time))
sales_dat$month<- month(as.POSIXct(sales_dat$Time))

sales_dat$day <- fct_inorder(sales_dat$day)
sales_dat<-sales_dat %>% filter(Name != "Missing", month != 4)
```

## Data Visualization


### Monthly 

```{r plot_data, fig.width=15, message=FALSE}
# Use this R-Chunk to plot & visualize your data!
 
 

sales_sum_Week <- sales_dat %>% 
  group_by(Name, week, day) %>% 
  summarise(Rev = sum(Amount))
sales_sum_Week$day <- case_when(
  sales_sum_Week$day == "Monday" ~ 0,
  sales_sum_Week$day == "Tuesday" ~ 1,
  sales_sum_Week$day == "Wednesday" ~ 2,
  sales_sum_Week$day == "Thursday" ~ 3,
  sales_sum_Week$day == "Friday" ~ 4,
  sales_sum_Week$day == "Saturday" ~ 5,
  sales_sum_Week$day == "Sunday" ~ 6)
sales_sum_Week$week_d <- sales_sum_Week$week + sales_sum_Week$day/7

sales_sum_Week_avg <- sales_sum_Week %>% group_by(Name, week) %>% summarise(Rev_avg = mean(Rev))

sales_sum <- sales_dat %>% 
  group_by(Name, day, hour) %>% 
  summarise(Rev = sum(Amount))
sales_sum_D <- sales_dat %>% 
  group_by(Name, day) %>% 
  summarise(Rev_sum = sum(Amount))
sales_sum_W <- sales_dat %>% 
  group_by(Name) %>% 
  summarise(Rev_sum = sum(Amount))
sales_sum_W$day = "Sunday"


sales_sum_M <- sales_dat %>% 
  group_by(Name, month) %>% 
  summarise(Rev = sum(Amount))

#levels(sales_sum$day)

#labels_week <- c("Monday \n 4","Tuesday","Wednesday", "Thursday", "Friday", "Saturday", "Sunday")   


# Monthly:

ggplot() +
  geom_col(aes(x=month, y=Rev, fill=Name), position="dodge" ,  data=sales_sum_M)+
  #geom_hline(aes(yintercept = Mean7day), bake10M)+
  geom_text(
    aes(label = paste(round(Rev,0)), x= month ,y = round(Rev,0)),vjust=1.2,
    position = position_dodge(),
    vjust = 0, sales_sum_M)+
  facet_wrap(~Name)+
  labs(fill = "Company", title= "Monthly Sales by company")+ 
  ylab("Revenue")+theme_bw()

sales_sum_M %>% 
    pivot_wider(names_from = month, values_from = Rev) %>% mutate(Total = `5` + `6` +`7`) %>% pander(caption= "Monthly total Sales by Company")

```

### Weekly

```{r fig.width=15, message=FALSE}
# Weekly/Monthly

week_list <- c(20,22,24,26,28,30)
ggplot(sales_sum_Week)+ 
   geom_line(aes(x =as.numeric(week_d), y= Rev, color=Name),size=1,alpha=0.25)+
   geom_line(aes(x =as.numeric(week), y= Rev_avg, color=Name),size=1, sales_sum_Week_avg)+
   labs(x="Week")+
  facet_wrap(~Name)+
  theme_bw()+
scale_x_discrete(limits = week_list)+
  labs(color= "Company", title= "Weekly average Sales by Company - Overlayed on Daily Sales")+ 
  ylab("Revenue")+
  xlab("Week")



# Weekly/Daily


ggplot(data=sales_sum) + 
    geom_raster(aes(day, hour , fill= Rev)) +
    theme_bw()+
  facet_wrap(~Name)+
    geom_text(
    aes(label = paste(Rev_sum), y=1 ,x= day),
    position = position_dodge(),
    vjust = 0, sales_sum_D)+
  geom_text(
    aes(label = paste("Total \n Sales: \n", Rev_sum), y=16 ,x= day),
    position = position_dodge(),
    vjust = 0, xjust= -2, sales_sum_W)+
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  geom_hline(yintercept = 8)+
  scale_fill_gradientn(colors= c( "red","yellow", "greenyellow","green","darkgreen"))+
  labs(title= "Daily Sales by hour and company", fill= "Sales")+ xlab("Day")+ ylab("Hour")+ theme_bw() #scale_x_discrete(labels= labels_week)

```

### Daily

```{r fig.width=15, message=FALSE}

# Daily
ggplot() +
  geom_col(aes(x=day, y=Rev_sum, fill=Name), position="dodge" ,  data=sales_sum_D)+
  #geom_hline(aes(yintercept = Mean7day), bake10M)+
  geom_text(
    aes(label = paste(round(Rev_sum,0)), x= day ,y = round(Rev_sum,0)),vjust=1.2,
    position = position_dodge(),
    vjust = 0, sales_sum_D)+
  geom_text(
    aes(label = paste("Total \n Sales: \n", round(Rev_sum,0)), y=6000 ,x= day),
    position = position_dodge(),
    vjust = 0, xjust=0, sales_sum_W)+
  facet_wrap(~Name)+
  labs(fill = "Episode Number", title= "Change in 7 day viewer over episode number by series")+ 
  ylab("7-Day Viewers (millions)")+theme_bw()

sales_sum_D %>%
    pivot_wider(names_from = day, values_from = Rev_sum) %>% mutate_all(~replace(., is.na(.), 0)) %>% mutate("Average Weekly"= case_when(`Sunday` > 0 ~ (`Monday` + `Tuesday` + `Wednesday` + `Thursday` + `Friday` + `Saturday` + `Sunday`)/7, `Sunday` <= 0  ~ (`Monday` + `Tuesday` + `Wednesday` + `Thursday` + `Friday` + `Saturday`)/6  )) %>% pander(caption= "Daily total Sales by Company")

```

## Conclusions

Based on the data as shown in the above visualization I feel that the best investment would be LeBelle. Shown in the average total weekly sales we see them fall just below HotDiggity but we can also observe in the monthyl sales they follow a left skewed distribution rather than the normal distribution of the other companies, without knowing what they sell it is difficult to infer but the other 5 companies may be affected by a different seasonality trend than LeBelle. This could mean LeBelle has a larger peak window for sales during the year. Additionally as can be seen in the hourly-weekly sales it would appear that based on POS data LeBelle is generally open for less time than many of these other companies, essentially only open for 6 days a week and closing sooner most nights than others but still managing to bring similar if not greater revenues than other companies analyzed.