---
title: "W10T1"
author: "Tyson Brost"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(readxl)
library(mosaic)
library(pander)
library(lubridate)
library(tidyquant)
#install.packages("tidyquant")
# BCC LPX JRV.V
# 
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
Friend_port <- data.frame()
stocklist  <- c("BCC", "LPX", "JRV.V") 
  for(i in stocklist){
    x <- tq_get(i, get = "stock.prices", from = " 2021-10-01")
    Friend_port <- rbind(x, Friend_port) 
}



My_port <- data.frame()
stocklist  <- c("TWLO", "AYX", "SPLK") 
  for(i in stocklist){
    x <- tq_get(i, get = "stock.prices", from = " 2021-10-01")
    My_port <- rbind(x, My_port) 
    }


```

## Background

_Place Task Background Here_

## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
Friend_port <- Friend_port %>%
    group_by(symbol) %>% 
    tq_transmute(select     = adjusted,
                 mutate_fun = dailyReturn,
                 type       = "log",
                 col_rename = "returns")

Friend_port1 <- Friend_port %>%
  group_by(symbol) %>%
  mutate(cum_ret = cumsum(returns))

Friend_port_sum <- Friend_port1 %>% 
  group_by(date) %>%
  summarise(returns = sum(returns))

My_port <- My_port %>%
    group_by(symbol) %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = dailyReturn,
                 type       = "log",
                 col_rename = "returns") 
My_port1 <- My_port %>%
  group_by(symbol) %>%
  mutate(cum_ret = cumsum(returns))
My_port_sum <- My_port1 %>% 
  group_by(date) %>%
  summarise(returns = sum(returns))


weights <- c(1/3, 1/3, 1/3)

port_friend <- Friend_port %>%
    tq_portfolio(assets_col   = symbol, 
                 returns_col  = returns, 
                 weights      = data.frame(asset.names = c("BCC", "LPX", "JRV.V"),
                weight = c(1/3, 1/3, 1/3)),
                 col_rename   = "Friend_Growth",
                 wealth.index = T)
port_my <- My_port %>%
    tq_portfolio(assets_col   = symbol, 
                 returns_col  = returns, 
                 weights      = data.frame(asset.names = c("TWLO", "AYX", "SPLK"),
                weight = c(1/3, 1/3, 1/3)),
                 col_rename   = "My_Growth",
                 wealth.index = T)
```

## Data Visualization

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!
ggplot(My_port1)+ 
   geom_line(aes(x =date, y= cum_ret, color=symbol),size=1)+
  facet_grid(rows=~symbol)+
  labs(color= "Stock", title= "Cumulative returns on my Portfolio by stock - Tech Industry")+ 
  ylab("Cumulative returns")+
  xlab("Date")

ggplot(My_port_sum)+ 
   geom_line(aes(x =date, y= cumsum(returns), color="Tech Industry"),size=1)+
  labs(title= "Cumulative returns on my Portfolio - Tech Industry")+ 
  ylab("Cumulative returns")+
  xlab("Date")+
  scale_color_manual(values = c("Tech Industry" = "darkred"))

ggplot(Friend_port1)+ 
   geom_line(aes(x =date, y= cum_ret, color=symbol),size=1)+
  facet_grid(rows=~symbol)+
  labs(color= "Stock", title= "Cumulative returns on my Friends Portfolio by stock - Mining and Forestry")+ 
  ylab("Cumulative returns")+
  xlab("Date")

ggplot(Friend_port_sum)+ 
   geom_line(aes(x =date, y= cumsum(returns), color= "Mining & Forestry"),size=1)+
  labs(title= "Cumulative returns on my Friends Portfolio - Mining and Forestry")+ 
  ylab("Cumulative returns")+
  xlab("Date")+
  scale_color_manual(values = c("Mining & Forestry" = "darkgreen"))


ggplot()+ 
   geom_line(aes(x =date, y= My_Growth, color="My Port"),size=1, data= port_my)+
  geom_line(aes(x =date, y= Friend_Growth, color="Friends Port"),size=1, data= port_friend)+
  labs(color= "Who's Portfolio?", title= "Comparion of Growth by Portfolio - Tech vs. Forestry/Mining")+ 
  ylab("Portfolio Growth")+
  xlab("Date")+
  scale_color_manual(values = c("My Port" = "darkred", "Friends Port" = "darkgreen"))
```

## Conclusions
As shown above, My hypothetical friends portfolio based in the forestry and mining industries performed significantly better than my own, based in tech/software companies. The two stocks helping the most where both Idaho based, one being Boise Cascade and the other Jervois (int. mining company w/ mines in Idaho), in comparison to Lousiana-Pacific. The one that hurt my portfolio the most was TWILIO compared to Splunk and Alterxy.