---
title: "My consulting Project"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


```{r, include=FALSE}
library(mosaic)
library(ResourceSelection) 
library(car)
library(tidyverse)
library(DT)
library(pander)
library(ggplot2)
library(readr)
library(stringr)
```

## R Markdown

```{r cars}
lab <- read_csv("C:/Users/user/OneDrive/Desktop/Stats stuff/325 book/Statistics-Notebook-master/CenterVisitReport (20).csv")
labmin <- lab %>%
  rename(Course = `Course`,
         Professor = `Professor`,
         Slot_type = `Slot type`,
         Students_class = `Student's class`,
         Student_type = `Student's type`,
         Visit_duration = `Visit duration (in minutes)`,
         DateTime = `Date (of the appointment/visit)`,
         Status = `Status (of the appointment/visit)`)

#labmin <- filter(lab, `Visit_duration` >= 1)

Slist <- c("MATH 221A", "MATH 221B", "MATH 221C", "MATH 325", "MATH 425", "MATH 326")
Mlist <- c("MATH 100A", "MATH 100B", "MATH 101", "MATH 108X", "MATH 109", "MATH 110X", "MATH 111", "MATH 112X", "MATH 113", "MATH 119", "MATH 191", "MATH 205", "MATH 214","MATH 215", "MATH 275", "MATH 280", "MATH 316", "MATH 330", "MATH 340", "MATH 341", "MATH 350")
labmin$DateTimeFor <- as.POSIXlt(labmin$DateTime, tz = "MST", format = "%m/%d/%Y %H:%M",
           tryFormats = c("%Y-%m-%d %H:%M:%OS",
                          "%Y/%m/%d %H:%M:%OS",
                          "%Y-%m-%d %H:%M",
                          "%m/%d/%Y %H:%M",
                          "%Y-%m-%d",
                          "%Y/%m/%d"))


labmin$Day_of_week <- weekdays(as.Date(labmin$DateTimeFor))
labmin$Dur_LT_one <- case_when(labmin$Visit_duration < 1 ~ 1, labmin$Visit_duration >= 1 ~ 0)
filteredstuff <- filter(labmin, (labmin$Course %in% Slist == TRUE),
                        (labmin$Dur_LT_one == 0))
MuStats <- mean(filteredstuff$Visit_duration)
filteredstuff <- filter(labmin, (labmin$Course %in% Mlist == TRUE),
                        (labmin$Dur_LT_one == 0))
MuNonStats <- mean(labmin$Visit_duration)
filteredstuff <- filter(labmin, (labmin$Course %in% Slist == FALSE & labmin$Course %in% Mlist == FALSE),
                        (labmin$Dur_LT_one == 0))
MuNonMath <- mean(labmin$Visit_duration)
#c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", 
#   "Friday", "Saturday")[as.POSIXlt(df$date)$wday + 1]
```


```{r}
library(tidyr)
library(stringr)

#labmindi <- filter(labmin, labmin$`Slot_type`== "DropIn")
#labmin <-labmin
#labmin <- drop_na(labmin$Course)
labmin <- labmin[!is.na(labmin$Course),]
labmin <- select(labmin, -X9)

labmin$nonmath <- case_when(labmin$`Course` %in% Slist == FALSE & labmin$`Course` %in% Mlist == FALSE ~ 1, labmin$`Course` %in% Slist == TRUE | labmin$`Course` %in% Mlist == TRUE ~ 0)
labmin$Check_In <- format(as.POSIXct(labmin$DateTimeFor), format = "%H:%M")
labmin$Date <- format(as.POSIXct(labmin$DateTimeFor), format = "%Y-%m-%d")
labmin$Day_of_week <- factor(labmin$Day_of_week)

labmin$`Visit_duration` <- case_when(labmin$`Course` %in% Slist == TRUE & (labmin$`Visit_duration` ==60 | labmin$`Visit_duration` < 1) ~ MuStats, 
                                     labmin$`Course` %in% Mlist == TRUE & (labmin$`Visit_duration` ==60 | labmin$`Visit_duration` < 1) ~ MuNonStats, 
                                     labmin$`Course` %in% Mlist == FALSE & labmin$`Course` %in% Slist == FALSE & (labmin$`Visit_duration` ==60 | labmin$`Visit_duration` < 1) ~ MuNonMath, 
                                     labmin$`Visit_duration` !=60 ~ labmin$`Visit_duration` )
#labmin$Check_In <-substr(labmin$Check_In, start = 1, stop = 2)
hour <- as.numeric(substr(labmin$Check_In, start = 0, stop = 2))
min <- (1.6666667 * as.numeric(substr(labmin$Check_In, start = 4, stop = 5)))/100
labmin$checkinNum <- hour + min


labmindp <-labmin
#find Q1, Q3, and interquartile range for values in column family
Q1 <- quantile(labmindp$Visit_duration, .25, na.rm = TRUE)
Q3 <- quantile(labmindp$Visit_duration, .75, na.rm = TRUE)
IQR <- IQR(labmindp$`Visit_duration`, na.rm = TRUE)


#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
LCD <- subset(labmindp, labmindp$`Visit_duration`> (Q1 - 1.5*IQR) & labmindp$`Visit_duration`< (Q3 + 1.5*IQR))



#LCD$stats <- case_when(LCD$`Course` %in% Slist == TRUE ~ 1, LCD$`Course` %in% Mlist == TRUE  ~ 0 )





LCDAVG <- LCD


LCDISD <- filter(LCD, LCD$`Course` %in% Slist == TRUE)
LCDNSD <- filter(LCD, LCD$`Course` %in% Mlist == TRUE)
LCDNMath <- filter(LCD, LCD$`Course` %in% Mlist == FALSE & LCD$`Course` %in% Slist == FALSE)

```

***
date <- LCDAVG$DateTime

#date <- str_replace_all(date, "/", "-")
#labmin %>% separate(labmin$Check_In, c("Hour", "Min"))

df <- x
            

#View(date)

start <- as.character(format(as.Date(min(df$date))),"yyyy-mm-dd") # Start 
end = as.character(format((max(df$date))),"yyyy-mm-dd") # End 
min = as.character(format(as.Date(min(df()$date))),"yyyy-mm-dd")
max = as.character(format(as.Date(max(df()$date))),"yyyy-mm-dd")
    
date<- as.character(LCDNSD$Date)
Checkin<- (LCDNSD$Check_In)
Course <- LCDNSD$Course
NonStats <- data.frame(date,Checkin,Course,stringsAsFactors = FALSE)
x <- NonStats



***


## Including Plots

### Overall sample

```{r pressure, echo=FALSE}


ggplot(LCDAVG, aes(x =as.numeric(LCDAVG$Check_In)))+
  geom_histogram()+
  facet_wrap(LCDAVG$Day_of_week)
```


### in stats

```{r}
ggplot(LCDISD, aes(x =as.numeric(LCDISD$Check_In)))+
  geom_histogram()+
  facet_wrap(LCDISD$Day_of_week)
```

### not in stats

```{r}
ggplot(LCDNSD, aes(x =as.numeric(LCDNSD$Check_In)))+
  geom_histogram()+
  facet_wrap(LCDNSD$Day_of_week)

#favstats(as.numeric(LCD$Check_In)~ LCD$Day_of_week)
```

```{r}
require(ggplot2)
require(reshape2)
#LCD
#set.seed(1)


#ggplot(melt(LCD), aes(value, fill = )) + geom_histogram(position = "dodge")

```


### Long Data

```{r}
library(reshape2)
LCDAVG$index <- 1:nrow(LCDAVG)

LCDLongCourse <- melt(LCDAVG,
  id.vars=c("index"),
  measure.vars=c("Course"),
  variable.name="Factor-Course",value.name="Course")
#INV_long$Year <- as.numeric(as.character(INV_long$Year))

LCDLongVisit_duration <- melt(LCDAVG,
  id.vars=c("index"),
  measure.vars=c("Visit_duration"),
  variable.name="Factor-Duration",value.name="Minutes")

LCDLongCheck_In <- melt(LCDAVG,
  id.vars=c("index"),
  measure.vars=c("Check_In"),
  variable.name="Factor-CheckIn",value.name="CheckIn")

LCDLongDay_of_week <- melt(LCDAVG,
  id.vars=c("index"),
  measure.vars=c("Day_of_week"),
  variable.name="Factor-Day",value.name="Day")

LCDLongDate <- melt(LCDAVG,
  id.vars=c("index"),
  measure.vars=c("Date"),
  variable.name="Factor-Date",value.name="Date")

LCDLongCourseDuration <- merge(LCDLongCourse,LCDLongVisit_duration,by="index")
LCDLongCheckInDayofWeek <- merge(LCDLongCheck_In,LCDLongDay_of_week,by="index")
LCDLongCoDuDa <- merge(LCDLongCourseDuration,LCDLongDate,by="index")
LCDLongMerge <- merge(LCDLongCoDuDa, LCDLongCheckInDayofWeek, by="index")

barplot(table(LCDLongMerge$Course, LCDLongMerge$Day), beside=TRUE, col=c("firebrick","snow1"), legend.text=TRUE, xlab="Course")
ggplot(mtcars, aes(x = factor(Day_of_week), fill = factor(Course), )
       )+ 
   geom_bar(position = "dodge", 
color = "black")+ 
   labs(x="Day_of_week")


library(data.table)

# get data
  data("LCDAVG")
LCDAVG_ND <- LCDAVG[-c(13)]
# transpose
  L_LCD <- transpose(LCDAVG_ND)

# get row and colnames in order
  colnames(L_LCD) <- rownames(LCDAVG_ND)

```



### Data 

```{r data}
LCDAVGs <- LCDAVG[sample(nrow(LCDAVG), 30), ]
```

```{r}
pander(LCDAVGs)
```
