---
title: "W8 Task 1"
author: "YOUR NAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(readxl)
library(mosaic)
library(pander)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
temp <- tempfile()
download.file("https://byuistats.github.io/M335/data/chipotle_reduced.csv",temp)
restDat <- read.csv(temp)
unlink(temp)
```

## Background

Read in the restaurant level data: https://byuistats.github.io/M335/data/chipotle_reduced.csv (Links to an external site.)
Here is a data dictionary: https://docs.safegraph.com/docs/monthly-patterns (Links to an external site.)

Create 2 functions. Both functions should take as input the character string that contains a count of visits by day of the week popularity_by_day. However, the 2 functions will differ in what they return:
Function 1 should return a dataframe (or tibble to me more precise) that contains the name of each day of the week in one column and the number of visits at that store in the other column

Function 2 should build on Function 1 and take it a step further. Namely, it should return the most popular day of the week (in terms of visits)

Verify your functions are working by calling them on the data for the following restaurants.
Restaurant Place Key

zzw-222@5vg-nwf-mp9

22c-222@5z5-3rs-hwk

zzw-223@5r8-fqv-xkf

## Data Wrangling

```{r}
#Functions

splitlist <- function(x){
x <- str_replace_all(x, ",", ":")
x <- str_replace_all(x, "\\{", "-")
x <- str_remove_all(x, "-")
x <- strsplit(x, ":", fixed = T)
}

build_table <- function(placekey1){
Filt <- select(filter(restDat, restDat$placekey == placekey1 ), `popularity_by_day`)
x <- splitlist(Filt)

result <- cbind(Filt[rep(1:nrow(Filt), lengths(x)), 1], content = unlist(x))

results <- data.frame(
  "day" = result[seq(1,length(result[,1]),2), "content"], 
  "Num_visitors" = result[seq(2,length(result[,1]),2), "content"])
results$day <- str_remove_all(results$day, "\"")
results$day <- str_replace_all(results$day, "[^[:alnum:]]", "") 
results$Num_visitors <- str_replace_all(results$Num_visitors, "[^[:alnum:]]", "") 
results %>% pander(caption= placekey1)
}


#placekey1 <-"zzw-222@5vg-nwf-mp9"

MostPop <- function(placekey1){
Filt <- select(filter(restDat, restDat$placekey == placekey1 ), `popularity_by_day`)
x <- splitlist(Filt)

result <- cbind(Filt[rep(1:nrow(Filt), lengths(x)), 1], content = unlist(x))

results <- data.frame(
  "day" = result[seq(1,length(result[,1]),2), "content"], 
  "Num_visitors" = result[seq(2,length(result[,1]),2), "content"])
results$day <- str_remove_all(results$day, "\"")
results$day <- str_replace_all(results$day, "[^[:alnum:]]", "")
results$Num_visitors <-str_replace_all(results$Num_visitors, "[^[:alnum:]]", "") 
M <- max(as.numeric(results$Num_visitors), data=results)
results <- select(filter(results, results$Num_visitors == M ), `day`)
results %>% pander(caption= i)
}

```

```{r tidy_data, eval=FALSE}
# Use this R-Chunk to clean & wrangle your data!
#does same thing as functions but in a loop
restlist <- list("zzw-222@5vg-nwf-mp9", "22c-222@5z5-3rs-hwk", "zzw-223@5r8-fqv-xkf")



for(i in restlist){
Filt <- select(filter(restDat, restDat$placekey == i ), `popularity_by_day`)
x <- splitlist(Filt)



result <- cbind(Filt[rep(1:nrow(Filt), lengths(x)), 1], content = unlist(x))

results <- data.frame(
  "day" = result[seq(1,length(result[,1]),2), "content"], 
  "Num_visitors" = result[seq(2,length(result[,1]),2), "content"])
results$day <- str_remove_all(results$day, "\"")
results$day <- str_replace_all(results$day, "[^[:alnum:]]", "") 
results$Num_visitors <- str_replace_all(results$Num_visitors, "[^[:alnum:]]", "") 
results %>% pander(caption= i)
}


```

## Data Visualization

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!

build_table("zzw-222@5vg-nwf-mp9")
build_table("22c-222@5z5-3rs-hwk")
build_table("zzw-223@5r8-fqv-xkf")

MostPop("zzw-222@5vg-nwf-mp9")
MostPop("22c-222@5z5-3rs-hwk")
MostPop("zzw-223@5r8-fqv-xkf")

```

## Conclusions
