---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(ggplot2)
library(readr)
food <- read_csv("../../Data/food.csv") #food.csv is in the Data folder...
```


----

<!-- Instructions:

 Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * GPA - numerical, self-reported current college GPA

  
  * weight - numerical, self-reported weight in lbs


  * gender - "Female", "Male"


  * breakfast - "Cereal", "Donut" 
      The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."


  * calories_chicken - guessing calories in chicken piadina from the options of:
     1 - 265 
     2 - 430 
     3 - 610 
     4 - 720
    (the variable shows the actual number of calories participants selected)


   * calories_day - Importance of consuming calories per day
     1 - i dont know how many calories i should consume 
     2 - it is not at all important 
     3 - it is moderately important 
     4 - it is very important


   * comfort_food - List 3-5 comfort foods that come to mind. 
     Open ended 


   * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
     Open ended


   * comfort_food_reasons_coded
     1 - stress
     2 - boredom
     3 - depression/sadness
     4 - hunger
     5 - laziness
     6 - cold weather
     7 - happiness 
     8 - watching tv
     9 - none 


    * cook - how often do you cook?
     1 - Every day 
     2 - A couple of times a week 
     3 - Whenever I can, but that is not very often  
     4 - I only help a little during holidays 
     5 - Never, I really do not know my way around a kitchen


    * cuisine - what type of cuisine did you eat growing up?
     1 - American
     2 - Mexican.Spanish
     3 - Korean/Asian
     4 - Indian
     5 - American inspired international dishes
     6 - other


    * diet_current - describe your current diet
     open ended

	
    * diet_current_coded
     1 - healthy/balanced/moderated/
     2 - unhealthy/cheap/too much/random/
     3 - the same thing over and over
     4 - unclear


    * drink - which picture do you associate with the word “drink”?
     1 - orange juice
     2 - soda 


    * eating_changes  - Describe your eating changes since the moment you got into college?
     Open ended 


    * eating_changes_coded
     1 - worse
     2 - better
     3 - the same
     4 - unclear


    * eating_changes_coded1
     1 - eat faster
     2 - bigger quantity
     3 - worse quality 
     4 - same food
     5 - healthier
     6 - unclear
     7 - drink coffee 
     8 - less food
     9 - more sweets
     10 - timing 
     11 - more carbs or snacking
     12 - drink more water
     13 - more variety


    * eating_out - frequency of eating out in a typical week 
     1 - Never 
     2 - 1-2 times 
     3 - 2-3 times 
     4 - 3-5 times 
     5 - every day


    * employment - do you work? 
     1 - yes full time 
     2 - yes part time 
     3 - no
     4  - other


    * ethnic_food - How likely to eat ethnic food 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * exercise - how often do you exercise in a regular week?
     1 - Everyday 
     2 - Twice or three times per week 
     3 - Once a week
     4 - Sometimes 
     5 - Never


    * father_education - 
     1 - less than high school 
     2 - high school degree 
     3 - some college degree 
     4 - college degree 
     5 - graduate degree 


    * father_profession - what is your father profession?
     Open ended


    * fav_cuisine - What is your favorite cuisine?
     Open ended


    * fav_cuisine_coded
     0-none
     1 - Italian/French/greek
     2 - Spanish/mexican
     3 - Arabic/Turkish
     4 - asian/chineses/thai/nepal
     5 - American
     6 - African 
     7 - Jamaican
     8 - indian


    * fav_food - was your favorite food cooked at home or store bought? 
     1 - cooked at home 
     2 - store bought 
     3 - both bought at store and cooked at home


    * food_childhood - what was your favorite childhood food?
     Open ended



    * fries - which of these pictures you associate with word fries? 
     1 - Mcdonald’s fries
     2 - home fries

 
    * fruit_day - How likely to eat fruit in a regular day 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * grade_level – 
     1 - freshman 
     2 -Sophomore 
     3 - Junior 
     4 - Senior
     
      
    * greek_food - How likely to eat greek food when available?
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 
     
      
    * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
    * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
      Open ended
      
      
    * ideal_diet – describe your ideal diet in 2-3 sentences
      Open ended
      
      
    * Ideal_diet_coded
      1 – portion control
      2 – adding veggies/eating healthier food/adding fruit	
      3 – balance
      4 – less sugar
      5 – home cooked/organic
      6 – current diet
      7 – more protein
      8 – unclear
      
      
     * income
      1 - less than $15,000 
      2 - $15,001 to $30,000 
      3 - $30,001 to $50,000 
      4 - $50,001 to $70,000 
      5 - $70,001 to $100,000 
      6 - higher than $100,000
      
     * indian_food – how likely are you to eat indian food when available
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * Italian_food – how likely are you to eat Italian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
     * marital_status
      1 -Single 
      2 - In a relationship 
      3 - Cohabiting 
      4 - Married 
      5 - Divorced 
      6 - Widowed
      
      
     * meals_dinner_friend – What would you serve to a friend for dinner?
      Open ended
      
      
     * mothers_education 
      1 - less than high school 
      2 - high school degree 
      3 - some college degree 
      4 - college degree 
      5 - graduate degree
      
      
     * mothers_profession – what is your mother’s profession? 
      
      
     * nutritional_check - checking nutritional values frequency 
      1 - never 
      2 - on certain products only 
      3 - very rarely 
      4 - on most products 
      5 - on everything
      
      
     * on_off_campus – living situation
      1 - On campus 
      2 - Rent out of campus 
      3 - Live with my parents and commute 
      4 - Own my own house
      
      
     * parents_cook - Approximately how many days a week did your parents cook? 
      1 - Almost everyday 
      2 - 2-3 times a week 
      3 - 1-2 times a week 
      4 - on holidays only 
      5 - never
      
      
     * pay_meal_out - How much would you pay for meal out? 
      1 - up to $5.00 
      2 - $5.01 to $10.00 
      3 - $10.01 to $20.00 
      4 - $20.01 to $30.00 
      5 - $30.01 to $40.00 
      6 - more than $40.01 
      
      
     * Persian_food - How likely to eat persian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * self_perception_weight - self perception of weight 
      6 - i dont think myself in these terms 
      5 - overweight 
      4 - slightly overweight 
      3 - just right 
      2 - very fit 
      1 - slim 
      
      
     * Which of the two pictures you associate with the word soup?
      1 – veggie soup
      2 – creamy soup
      
      
     * sports - sports – do you do any sporting activity?
      1 - Yes 
      2 - No 
      99 – no answer
      
      
     * thai_food - How likely to eat thai food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
      1 - 580 
      2 - 725 
      3 - 940 
      4 - 1165
      
     * turkey_calories - Can you guess how many calories are in the foods shown below?
     (Panera Bread Roasted Turkey and Avocado BLT)
      1 - 345 
      2 - 500 
      3 - 690 
      4 - 850
      
      
     * type_sports – what type of sports are you involved?
      Open-ended
      
     * veggies_day - How likely to eat veggies in a day? 
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4- likely 
      5 - very likely
      
     * vitamins – do you take any supplements or vitamins?
      1 – yes
      2 – no
      
     * waffle_calories - guessing calories in waffle potato sandwhich 
      1 - 575 
      2 - 760 
      3 - 900 
      4 - 1315


Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:

[Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->




<!--End your comments to your critiquers before this line.-->

----

<!--Begin writing your Analysis after this point.-->

## Background
I have always heard that a healthy diet will help you perform better in school so I wanted to compare how much the likelihood of a student eating Fruits or Veggies in any given day affects GPA. 
The data that I will be using to test this hypothesis comes from a survey of college students. As the current data set doesn't contain a factor of propensity for veggies or fruits combined, just one or the other, I have decided to create a third factor that is representative of them both. This is done by combining the two 5 level factors into one 25 level factor with a new level for each possible combination.

Below is my attempt to explain the process by which I created the third factor. X axis is propensity to eat fruit and the Y axis is the propensity to each veggies, because I felt a propensity of 2 for veggies and a propensity of 1 for fruit was different than a propensity of 1 for veggies and a propensity of 2 for fruit, I didn't simply just take the product of x and y but rather a sum of the two levels and let veggies take the larger of the numbers when a tie was present.


 


+-----------------+----+----+----+----+----+
| Propensities    |    |    |    |    |    |
| (C:Fruit,R:Veg) | 1  | 2  |  3 |  4 |  5 |
+=================+====+====+====+====+====+
| 1               | 1  |  2 | 4  | 7  | 11 |
+-----------------+----+----+----+----+----+
| 2               | 3  | 5  | 8  | 12 | 16 |
+-----------------+----+----+----+----+----+
| 3               | 6  |  9 | 13 | 17 | 20 |
+-----------------+----+----+----+----+----+
| 4               | 10 | 14 | 18 | 21 | 23 |
+-----------------+----+----+----+----+----+
| 5               | 15 | 19 | 22 | 24 | 25 |
+-----------------+----+----+----+----+----+

```{r message=FALSE, warning=FALSE}
food[68, 2] = "144"
food[4, 2] = "240"

food2 <- food[-c(3,62,105),]
library(tidyr)
food2 <- food2 %>% drop_na(GPA, weight)
#Clean Data, had hanging "lbs", Said "Not sure, guess 240", and "I will not awnser"
```

                        
```{r}
#Create 3rd Factor of Fruit/Veggie interaction
food2$`Fruit-Veggie` <- case_when(food2$fruit_day == 1 & food2$veggies_day == 1 ~ 1, 
                                  food2$fruit_day == 1 & food2$veggies_day == 2 ~ 2,
                                  food2$fruit_day == 1 & food2$veggies_day == 3 ~ 6,
                                  food2$fruit_day == 1 & food2$veggies_day == 4 ~ 10,
                                  food2$fruit_day == 1 & food2$veggies_day == 5 ~ 15,
                                  food2$fruit_day == 2 & food2$veggies_day == 1 ~ 2,
                                  food2$fruit_day == 2 & food2$veggies_day == 2 ~ 5,
                                  food2$fruit_day == 2 & food2$veggies_day == 3 ~ 9,
                                  food2$fruit_day == 2 & food2$veggies_day == 4 ~ 14,
                                  food2$fruit_day == 2 & food2$veggies_day == 5 ~ 19,
                                  food2$fruit_day == 3 & food2$veggies_day == 1 ~ 4,
                                  food2$fruit_day == 3 & food2$veggies_day == 2 ~ 8,
                                  food2$fruit_day == 3 & food2$veggies_day == 3 ~ 13,
                                  food2$fruit_day == 3 & food2$veggies_day == 4 ~ 18,
                                  food2$fruit_day == 3 & food2$veggies_day == 5 ~ 22,
                                  food2$fruit_day == 4 & food2$veggies_day == 1 ~ 7,
                                  food2$fruit_day == 4 & food2$veggies_day == 2 ~ 12,
                                  food2$fruit_day == 4 & food2$veggies_day == 3 ~ 17,
                                  food2$fruit_day == 4 & food2$veggies_day == 4 ~ 21,
                                  food2$fruit_day == 4 & food2$veggies_day == 5 ~ 24,
                                  food2$fruit_day == 5 & food2$veggies_day == 1 ~ 11,
                                  food2$fruit_day == 5 & food2$veggies_day == 2 ~ 16,
                                  food2$fruit_day == 5 & food2$veggies_day == 3 ~ 20,
                                  food2$fruit_day == 5 & food2$veggies_day == 4 ~ 23,
                                  food2$fruit_day == 5 & food2$veggies_day == 5 ~ 25
                                  )
```




```{r}
#convert factors to numeric
food2$fruit_C <- as.character(food2$fruit_day)
food2$fruit_N <- as.numeric(food2$fruit_C)

food2$veg_C <- as.character(food2$veggies_day)
food2$veg_N <- as.numeric(food2$veg_C)

```                        




It is interesting to note that as shown below, there appears to possibly be some level of positive correlation between propensity to consume fruit and propensity to consume veggies. As the groups to the outer extremes in our 1st and 4th quadrants are much smaller if present at all. We only have 17 of the 25 possible level combinations as a possible result of this relationship. We can also see that low propensities for either fruit or veggies are relatively uncommon and the majority of the data sits with at least a 3 or higher in at least either fruit or veggie propensity, if not in both.

```{r}
#plot
ggplot(food2, aes( fruit_N , veg_N))+
  geom_count(aes(color=`Fruit-Veggie`))+
  labs(title="Fruit-Veggie Propensity groups", x="Fruit Propensity", y="Veggie Propensity", legend = "Count")
```
----

## Hypothesis
As stated before I will be looking to see if any of the groups differs from the others in respect to GPA distribution based on Propensity to eat Fruits-Veggies during a given day.
$$
  H_0: \text{ All samples (Propensity Groups) represent a sample of data taken from the same population (GPA's). } 
$$

$$
  H_a: \text{ At least one distribution is stochastically different than the others. } 
$$
A significance level of 0.05 will be used.

In other words, I am testing to see if any one of the propensity groups has a GPA that is in general different from the others.

----

## Analysis

Below is a boxplot showing the distributions of the 17 different levels from our new factor that we had Data for.
```{r message=FALSE}
ggplot(food2, aes(x=reorder(as.character(`Fruit-Veggie`), `Fruit-Veggie`), y=as.numeric(GPA), fill=`Fruit-Veggie`))+
   geom_boxplot()+
  labs(title="GPA grouped by Propensity to eat Fruit/Veggies", x="Fruit-Veggie propensity", y="GPA")+
  coord_cartesian(ylim = c(0, 4.5))

```

#### GPA Table
##### Data aggregated as no  significant variation is present.

```{r message=FALSE, warning=FALSE}
 favstats(food2$GPA)[-9] %>% 
    pander(caption = "GPA")

```


## K-Test

Here are the results of our Kruskal-Wallis test:
```{r message=FALSE, warning=FALSE}
kruskal.test(GPA ~ `Fruit-Veggie`, data = food2) %>%
pander()
```

----
## Conclusion

As shown we can see with a P-Value of 0.606 it is highly unlikely that any significant relationship with GPA and propensity to consume fruits or veggies exists, we fail to reject the null. As far as what the relationship for what students ACTUALLY eat we do not know, but at least for what students say they are likely to eat during a given day there does not appear to be any effect on average GPA.
