---
title: "W5T1"
author: "Tyson Brost"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(readr)
library(stringr)
library(reshape2)
library(ggplot2)
devtools::install_github('Ather-Energy/ggTimeSeries')
library(ggTimeSeries)
library(mosaic)
#install.packages("mosaic")
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
devtools::install_github("drsimonj/ourworldindata")
#install.packages("devtools")
library(devtools)
childMortality <- ourworldindata::child_mortality
```

## Background

Take notes on your reading of the specified ‘R for Data Science’ chapter in the README.md or in a ‘.R’ script in the class task folder.
Explore the world data on child_mortality and create an interesting graphic that highlights this dataset.
Use devtools::install_github("drsimonj/ourworldindata") to get a larger dataset on child_mortality. This will require install.packages("devtools").
Create a graphic that you could use in a presentation to summarize world child mortality.
Include your plots in an .Rmd file with short paragraph describing your plots.
Knit your .Rmd and push the files it generates to your GitHub repo.

## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
childMortality$continent <- as.factor(childMortality$continent)
childMortality$country <- as.factor(childMortality$country)

childMortality <- filter(childMortality, year >= 1900)
childMortality <- childMortality %>% drop_na(child_mort)
childMortality <- childMortality %>% drop_na(continent)
#means <- favstats(child_mort ~ continent, data=childMortality)
means <- childMortality %>% 
        group_by(continent, year) %>% 
        summarise(childmortM = mean(child_mort))


#I need to learn how to do a dynamic case when so I can run this by year
#childMortalityContM <- case_when(childMortality$continent == "Africa" ~ 270.1261,(childMortality$continent == "Asia" ~ 270.1261)
```

## Data Visualization

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!
ggplot(data=means, ) +
    geom_area(aes(x=year, y=childmortM, fill= continent)) +
    theme_bw() +
    facet_wrap(~continent) +
    theme(
      panel.spacing = unit(0.5, "lines"),
      axis.ticks.x=element_blank()
    )+ 
  labs(fill = "Continent", title= "Average Child mortality by Continent since 1900")+ 
  xlab("Year") + ylab("Child Mortality (Avg)")

#ggplot(childMortality, aes(x = year, y = mean(child_mort), group = continent, fill = continent))+
  #geom_area()+
# facet_wrap(~continent)+
 #stat_steamgraph()
```

## Conclusions
