---
title: "Map of Campus in r"
author: "Tyson Brost"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 8
    fig_width: 6
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(leaflet)

```

```{r load_data}
# Use this R-Chunk to import all your datasets!
BYUI <- data.frame(Building = c("STC","Austin","BYU-I Center","Hart","snow","Romney","Spori","Kirkham", "Clarke", "Mckay", "Smith", "Manwaring", "Kimball", "Taylor", "Hinckley", "Ricks", "Benson"), Lat = c(43.8144,43.81582909024388,43.81855179894589,43.81957363416934,43.82109432348679,43.820281522138444, 43.820916282233114, 43.82110980530984, 43.82008025533344, 43.819499674209986, 43.81936033390029, 43.81821463458064, 43.81703020609232, 43.81693730874667, 43.81573737168207, 43.814916755695485, 43.815381256695574 ), Long = c(-111.78459,-111.78448729257276,-111.78459523261063,-111.784820538147,-111.78355920298482,-111.78302276123155, -111.78255069248866, -111.78159582616784, -111.78179967403408, -111.78240048879773,-111.78144562247691, -111.78221809860163, -111.78145635131197,-111.78248631947825, -111.77987921247086, -111.78131687636964, -111.78314077833078), MainSubjects = c("Computer Science","Engineering","Devo","Recreation","Performance","Physical Sciences","Art","Sculpting?", "Nursing/Family Science", "Library", "Economics/Business", "Dance/University services", "University services", "Religion", "Languages", "Math/Psychology", "Plant/Animal Science"))
   

```

## Background

_Place Task Background Here_

## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
getColor <- function(BYUI) {
  sapply(BYUI$Building, function(Building) {
  if(Building %in% c("STC","Austin","Romney", "Smith", "Ricks", "Benson")) {
    "lightblue"
  } else if(Building %in% c("BYU-I Center","Hart", "Mckay","Manwaring", "Kimball")) {
    "green"
  } else {
    "firebrick"
  } })
}
colors <- getColor(BYUI)
colors <- unname(colors)
icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = colors
)

```

## Data Visualization

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!
m <- leaflet(data = BYUI) %>% setView(lng = -111.784, lat = 43.818, zoom = 15.5) %>% 
  addTiles() %>%
  addAwesomeMarkers(~Long, ~Lat, icon=icons, popup = ~as.character(MainSubjects), label=~as.character(Building))
m

```

## Conclusions
I pulled the Long and Lat data from google maps and built a quick DF to store the data. Then I used a couple functions from a leaflet I did back in 325 to add colors based on sublists of the building names, the rough categories are; blue -> STEM, green -> University services/ETC, red -> Arts
the Building names appears on hover, click shows the primary subject taught there.
